---
title: "Stat 3202 Lab 10"
author: "Nathan Johnson.9254"
date: "2023-04-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Guided Example 1:**
```{r}
library(readr)
vanillin <- c(38,46,41,36,42,40,34,41,33,
46,41,39,44,45,41,37,39,36,
40,43,37,41,39,36,47,36,41,
38,38,38,33,39,41,36,38,44,39)
mean(vanillin)
```
1. Confidence interval
Test $H_0: \mu = 42$ vs. $H_1: \mu \neq 42$ using a 99% confidence interval for $\mu$.

```{r}
alpha = 0.01
n = length(vanillin)
xbar = mean(vanillin)
sigma = 3.52

z = qnorm(1-alpha/2)

CIlower = xbar - z * sigma/sqrt(n)
CIupper = xbar + z * sigma/sqrt(n)

c(CIlower, CIupper)
```
We are 99% confident that the true mean vanillin level of this batch is between 38.05 and 41.03 ppt. Since 42 is not contained within this confidence interval, we reject the null hypothesis, $H_0$.

We conclude that the true mean vanillin level is statistically significantly different from 42.

2. Rejection region
We are testing the hypotheses $H_0: \mu = 42$ vs. $H_1: \mu \neq 42$.
$$Z = \frac{\bar x - 42}{\sigma / \sqrt n}$$
Under the null hypotheses, the test statistic Z follows the distribution $Z \sim N(0,1)$.

We reject the null hypothesis $H_0$ when we observe our test statistic $Z_{obs}$ to be either too high or too low.

```{r}
alpha = 0.01

rrlower = qnorm(alpha/2)
rrupper = qnorm(1-alpha/2)
c(rrlower, rrupper)
```

The rejection region is $(-\infty, -2.58]$ and $[2.58, \infty)$, and we would reject the null hypothesis when our observed test statistic falls within either of these regions.

```{r}
Z.obs = (xbar - 42) / (sigma / sqrt(n))
Z.obs
```
We observe $Z_{obs} = -4.25$. Since this falls within the rejection region, we reject the null hypothesis $H_0$. And we conclude that the $\mu \neq 42$. That is the true mean vanillin is statistically significantly different from 42 ppt.

3. Verify with Z Test
```{r}
library(DescTools)

ZTest(vanillin, alternative = "two.sided", mu = 42, sd_pop = 3.52, conf.level = 1-alpha)
```

**Problem 2:**
```{r}
library(readr)
vanillin <- c(38,46,41,36,42,40,34,41,33,
46,41,39,44,45,41,37,39,36,
40,43,37,41,39,36,47,36,41,
38,38,38,33,39,41,36,38,44,39)
mean(vanillin)
```
1. Confidence interval
Test $H_0: \mu = 42$ vs. $H_1: \mu \neq 42$ using a 99% confidence interval for $\mu$.

```{r}
alpha = 0.01
n = length(vanillin)
xbar = mean(vanillin)
sigma = sd(vanillin)

t = qt(1-alpha/2, n - 1)

CIlower = xbar - t * sigma/sqrt(n)
CIupper = xbar + t * sigma/sqrt(n)

c(CIlower, CIupper)
```
We are 99% confident that the true mean vanillin level of this batch is between 37.96 and 41.12 ppt. Since 42 is not contained within this confidence interval, we reject the null hypothesis, $H_0$.

We conclude that the true mean vanillin level is statistically significantly different from 42.

2. Rejection region
We are testing the hypotheses $H_0: \mu = 42$ vs. $H_1: \mu \neq 42$.
$$t = \frac{\bar x - 42}{s / \sqrt n}$$
Under the null hypotheses, the test statistic t follows the distribution $t \sim N(0,1)$.

We reject the null hypothesis $H_0$ when we observe our test statistic $t_{obs}$ to be either too high or too low.

```{r}
alpha = 0.01

rrlower = qt(alpha/2, n - 1)
rrupper = qt(1-alpha/2, n - 1)
c(rrlower, rrupper)
```

The rejection region is $(-\infty, -2.72]$ and $[2.72, \infty)$, and we would reject the null hypothesis when our observed test statistic falls within either of these regions.

```{r}
t.obs = (xbar - 42) / (sigma / sqrt(n))
t.obs
```
We observe $t_{obs} = -4.24$. Since this falls within the rejection region, we reject the null hypothesis $H_0$. And we conclude that the $\mu \neq 42$. That is the true mean vanillin is statistically significantly different from 42 ppt.

3. Verify with t Test
```{r}
library(DescTools)

t.test(vanillin, alternative = "two.sided", mu = 42, sd_pop = sigma, conf.level = 1-alpha)
```
**Problem 3:**

a. We would use a z test statistic for this hypothesis.
b. Z distribution. 
$$Z_{obs} = \frac{\bar x - 8}{1 / \sqrt{30}} \sim^{H_0}N(0,1)$$.
```{r}
alpha = 0.05
zlower = qnorm(alpha/2)
zupper = qnorm(1-alpha/2)
c(zlower, zupper)
```
The rejection region is $(-\infty, -1.96], [1.96, \infty).$

c & d & e.
```{r}
set.seed(1)
n = 30
mu = 8
sigma = 1

zstat = c()
rrcoverage = c()

for(i in 1:20000) {
  samp = rnorm(n, mu, sigma)
  
  zstat[i] = (mean(samp) - mu) / (sigma / sqrt(n))
  
  rrcoverage[i] = zstat[i] <= zlower || zstat[i] >= zupper
}
mean(rrcoverage)

hist(zstat, col = "azure", xlab = "Values", main = "Histogram of Zobs ~ N(0,1)", breaks = 100)
```
d. My answer is 0.05025 which means 5% of the samples fall within the rejection region, thus is a 95% confidence interval. We are 95% confident that $Z_{obs}$ is captured in the confidence interval.

e. As can be seen, the histogram of $Z_obs$ does simulate a normal distribution.

**Problem 4:**

a. We would use a $\chi ^2$ test statistic for this hypothesis.
b. Chi-square distribution. 
$$\chi^2_{obs} = (19-1)s^2/1 \sim^{H_0}\chi^2_{19-1}$$.
```{r}
alpha = 0.01
n = 19

chiupper = qchisq(1-alpha, n - 1)
chiupper
```
The rejection region is $[34.81, \infty).$

c & d & e.
```{r}
set.seed(1)
mu = 8
sigma = 1

chistat = c()
rrcoverage = c()

for(i in 1:20000) {
  samp = rnorm(n, mu, sigma)
  var = var(samp)
  chistat[i] = (n - 1)*var
  
  rrcoverage[i] = chiupper <= chistat[i]
}
mean(rrcoverage)

hist(chistat, col = "azure", xlab = "Values", main = "Histogram of Xobs", breaks = 100)
```

d. My answer is 0.0092 which means roughly 1% of the samples fall within the rejection region, thus is a 99% confidence interval. We are 99% confident that $\chi^2_{obs}$ is captured in the confidence interval.

e. As can be seen, the histogram of $\chi^2_{obs}$ does simulate a chi-squared distribution.
