---
title: "Stat 3202 Lab 11"
author: "Nathan Johnson.9254"
date: "2023-04-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Guided Example 1:**

```{r}
library(readr)
library(tidyverse)
batch1 <- c(38,46,41,36,42,40,34,41,33,
46,41,39,44,45,41,37,39,36,
40,43,37,41,39,36,47,36,41,
38,38,38,33,39,41,36,38,44,39)
batch2=c(42, 37, 40, 39 ,45, 31, 45, 36, 37, 38,
41, 35, 46, 41, 44, 37, 38, 39, 38, 38,
42, 44,45, 35, 35, 42, 36, 40)
```

**Part a:**
Our test statistic is:
$$Z = \frac{(\bar x_1 - \bar x_2) - 0}{\sqrt{\frac{\sigma^2_1}{n_1}+\frac{\sigma^2_2}{n_2}}}$$

**Part b:**
Under the null hypothesis, the sampling distribution of this test statistic is $Z \sim N(0,1)$.

**Part c:**
```{r}
xbar1 = mean(batch1)
xbar2 = mean(batch2)

n1 = length(batch1)
n2 = length(batch2)

sigma21 = 3.5^2
sigma22 = 3.5^2

Z.obs = (xbar1 - xbar2) / sqrt(sigma21/n1 + sigma22/n2)
Z.obs
```

Observed test statistic is $Z_{obs} = 0.0462$.

**Part d:**
```{r}
alpha = 0.01

rr1 = qnorm(alpha/2)
rr2 = qnorm(1-alpha/2)

rr1
rr2

c(-Inf, rr1)
c(rr2, Inf)
```

Our rejection regions are $(-\infty, -2.58], [2.58, \infty)$.

We don't reject the null hypothesis because $Z_{obs} =  0.0462$ is not contained in the rejection region and so we fail to reject the null hypothesis.

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part e:**

We will construct a 99% 2-sided confidence interval for $\mu_1 - \mu_2$.

$$(\bar x_1 - \bar x_2) \pm z_{1-\alpha/2}\sqrt{\frac{\sigma^2_1}{n_1}+\frac{\sigma^2_2}{n_2}}$$
```{r}
z.quant = qnorm(1-alpha/2)

CIlower = (xbar1 - xbar2) - z.quant * sqrt(sigma21/n1 + sigma22/n2)
CIupper = (xbar1 - xbar2) + z.quant * sqrt(sigma21/n1 + sigma22/n2)

c(CIlower, CIupper)
```

We are 99% confident that the true difference in mean vanillin levels between batches 1 and 2 is between -2.22 and 2.30 ppt.

Since 0 is contained within our confidence interval, we fail to reject the null hypothesis. 

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part f:**

$$p = P(Z > |Z_{obs}|) * 2$$

```{r}
Z.obs

p = pnorm(Z.obs, lower.tail = FALSE) * 2
p
```

Since our p-value wasn't less than $\alpha = 0.01$, we fail to reject the null hypothesis.

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part g:**
```{r}
library(DescTools)

ZTest(batch1, batch2, alternative = "two.sided", mu = 0, sd_pop = 3.5, conf.level = 1-alpha)
```

**Problem 2:**

To see if population standard deviations are equal or not, we must test if $\frac{\sigma_1}{\sigma_2} = 1$. So we need a confidence interval, using the F distribution, of the ratio of the variances and see if it contains 1. If it does, then we can say that the population standard deviations are equal. If it does not, then standard deviations are not equal.

```{r}
sigma21 = var(batch1)
sigma22 = var(batch2)

n1 = length(batch1)
n2 = length(batch2)

alpha = 0.01

Flower = qf(alpha/2, n1 - 1, n2 - 2)
Fupper = qf(1-alpha/2, n1 - 1, n2 - 2)

CIlower = (sigma21/sigma22)/Fupper
CIupper = (sigma21/sigma22)/Flower

c(CIlower, CIupper)
```

Since the confidence interval contains 1, we will have to repeat question 1 assuming the population standard deviations are unknown, but equal. 

**Part a:**
We will be using a t statistic.

$$t = \frac{(\bar x_1 - \bar x_2) - 0}{s_p\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$$ where $$s_p = \frac{(n_1 - 1)s^2_1 + (n_2-1)s^2_2}{n_1+n_2-2}$$

**Part b:**
Under the null hypothesis, the sampling distribution of this test statistic is $t \sim t_{n_1+n_2-2}$.

**Part c:**
```{r}
xbar1 = mean(batch1)
xbar2 = mean(batch2)

n1 = length(batch1)
n2 = length(batch2)

sigma21 = var(batch1)
sigma22 = var(batch2)

spooled = sqrt(((n1 - 1)*sigma21 + (n2 - 1)*sigma22)/(n1+n2-2))

t.obs = (xbar1 - xbar2) / (spooled * sqrt((1/n1) + (1/n2)))
t.obs
```

Observed test statistic is $t_{obs} = 0.0446$.

**Part d:**
```{r}
alpha = 0.01

rr1 = qt(alpha/2, n1 + n2 - 2)
rr2 = qt(1-alpha/2, n1 + n2 -2)

rr1
rr2

c(-Inf, rr1)
c(rr2, Inf)
```
Our rejection regions are $(-\infty, -2.66], [2.66, \infty)$.

We don't reject the null hypothesis because $t_{obs} =  0.0446$ is not contained in the rejection region and so we fail to reject the null hypothesis.

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part e:**

We will construct a 99% 2-sided confidence interval for $\mu_1 - \mu_2$.

$$(\bar x_1 - \bar x_2) \pm t_{(1-\alpha/2,n_1+n_2-2)s_p\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$$
```{r}
t.quant = qt(1-alpha/2, n1 + n2 - 2)

spooled = sqrt(((n1 - 1)*sigma21 + (n2 - 1)*sigma22)/(n1+n2-2))

CIlower = (xbar1 - xbar2) - (t.quant * spooled * sqrt((1/n1) + (1/n2)))
CIupper = (xbar1 - xbar2) + (t.quant * spooled * sqrt((1/n1) + (1/n2)))

c(CIlower, CIupper)
```

We are 99% confident that the true difference in mean vanillin levels between batches 1 and 2 is between -2.37 and 2.46 ppt.

Since 0 is contained within our confidence interval, we fail to reject the null hypothesis. 

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part f:**
$$p = P(t > |t_{obs}|) * 2$$

```{r}
Z.obs

p = pnorm(t.obs, lower.tail = FALSE) * 2
p
```

Since our p-value wasn't less than $\alpha = 0.01$, we fail to reject the null hypothesis.

We conclude that there is no statistically significant difference in vanillin levels between batch 1 and batch 2.

**Part g:**
```{r}
library(DescTools)

t.test(batch1, batch2, alternative = "two.sided", mu = 0, var.equal = TRUE, conf.level = 1-alpha)
```


