---
title: "Stat 3202 Lab 3"
author: "Nathan Johnson.9254"
date: "2023-01-23"
output: pdf_document
---

```{r}
tidy.opts=list(width.cutoff=60, tidy=TRUE)
```
**Problem 1:**

CLT: Normalize a particular distribution
$\bar{X}$ has approximately mean $\mu$ and $\frac{\sigma^2}{n}$

That is: $\frac{\bar{X} - \mu}{\frac{\sigma}{n}} \overset{approx}{\sim} N(0, 1)$

**Part A:**
```{r}
set.seed(1)

mc = 10000
n = 4
xbar = c()

for (i in 1:mc) {
  sample = rgamma(n, 2, 0.5)
  xbar[i] = mean(sample)
}

hist(xbar, main = "Sampling Distribution of Sample Mean for n = 4", xlab = "Sample Mean", col = "azure")
```

This sampling distribution is slightly right skewed, so it is not normally distributed.

**Part B:**
```{r}
set.seed(1)

mc = 50000
n = 4
xbar = c()

for (i in 1:mc) {
  sample = rgamma(n, 2, 0.5)
  xbar[i] = mean(sample)
}

hist(xbar, main = "Sampling Distribution of Sample Mean for n = 4", xlab = "Sample Mean", col = "azure")
```

No, it doesn't change because all was changed was the simulations, the number of samples was not changed.

**Part C:**
```{r}
set.seed(1)

mc = 10000
n = 80
xbar = c()

for (i in 1:mc) {
  sample = rgamma(n, 2, 0.5)
  xbar[i] = mean(sample)
}

hist(xbar, main = "Sampling Distribution of Sample Mean for n = 80", xlab = "Sample Mean", col = "azure")
```

Yes, the sample was changed.

**Part D:**

The true mean of $\bar{X}$ is 4 and the variance is 8 as per the gamma distribution. 

```{r}
mean(xbar)
var(xbar)*80
```
The mean is slightly under 4 (3.996224) and the variance is also slightly under 8 (7.976128).

Due to Central Limit Theorem, $\bar{X}$ approximately has the distribution $N(4, 8/80)$.

**Part E:**

```{r}
set.seed(1)

mc = 10000
n = 50
xbar = c()

for (i in 1:mc) {
  sample = rcauchy(n)
  xbar[i] = mean(sample)
}

hist(xbar, main = "Sampling Distribution of Sample Mean for n = 50", xlab = "Sample Mean", col = "azure")
```


**Problem 2:**

**Part A:**
```{r}
set.seed(1)

pois = rpois(5000, 7)
barplot(table(pois), main = "Sampling Distribution of Poisson Distrbition", xlab = "Value", ylab = "frequency")
```

**Part B:**
```{r}
set.seed(1)

pois = rpois(5000, 7)
mean = mean(pois)
mean
var = var(pois)
var
```

Mean is closer to the true value of $\lambda$. 

**Part C:**
I believe we would choose the sample mean. The sample variance can be changed more than the mean will be by outliers.

**Problem 3:**

**Part A:**
As per the random variables:
$E(A) = 6, V(A) = 4, E(B) = 5, V(B) = 10, E(C) = \frac{1}{10}, V(C) = \frac{1}{100}$
To find the expectation of X, we have to do $E(X) = E(A+2B-3C) = E(A) + 2E(B) - 3E(C) = 6 + 2*5 - 3*\frac{1}{10} = 15.7$
To find the variance of X, we have to do $V(X) = V(A+2B-3C) = V(A) + 4V(B) + 9V(C) = 4 + 4*10 + 9*\frac{1}{100}=44.09$

**Part B:**
```{r}
set.seed(1)
n=5000

A=rnorm(n,6,2)
B=rchisq(n,5)
C=rexp(n,10)
X=A+2*B-3*C
EX=mean(X)
VX=var(X)

hist(X,main="Histogram of X", xlab = "Values")
EX
VX
```
The expectation of X is 15.75409 which is 0.05409 above the true mean and the variance of X is 44.63352 which is 0.543532 above the variance.

**Part C:**
```{r}
set.seed(1)
A=pnorm(3.5,6,2,lower.tail=TRUE)
A
```
The probability that an observation from distribution A is less than 3.5 is 0.1056498.

**Part D:**
```{r}
set.seed(1)
A=rnorm(5000,6,2)
lessThan=round(A<3.5)
lessThanMean=mean(lessThan)
lessThanMean
```
The true probability is 0.1056498 while the given probability is 0.1128 so they are pretty close. They aren't identical, but with greater and greater values of n, the distribution of A should get closer and closer to the true probability.

**Part E:**
```{r}
set.seed(1)
greaterThan=round(X>10)
greaterThanMean=mean(greaterThan)
greaterThanMean
```
The approximate probability of observations of X greater than 10 is 0.8082.