---
title: "Stat 3202 Lab 6"
author: "Nathan Johnson.9254"
date: "2023-02-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Problem 1:**
**Part a:**
$N(\mu,\frac{\sigma^2}{n})$.

**Part b:**
```{r}
set.seed(1)

mc = 50000

n = 40
mu = 20
x = sqrt(4)

storage = c()

for(i in 1:mc) {
  samp = rnorm(n, mu, x)
  storage[i] = mean(samp)
}

hist(storage, col = "blue", breaks = 50)
```
$N(20,\frac{2}{\sqrt{40}})$.

**Part c:**
```{r}
set.seed(1)

mc = 50000

n = 4
mu = 20
x = sqrt(4)

storage = c()

for(i in 1:mc) {
  samp = rnorm(n, mu, x)
  storage[i] = mean(samp)
}

hist(storage, col = "blue", breaks = 50)
```
$N(20,\frac{2}{\sqrt{4}})$.

**Part d:**
The sample sizes didn't impact the mean, but did impact the variation of the sampling distribution of $\bar{x}$.

**Part e:**
```{r}
set.seed(1)

mc = 50000

n = 9
mu = 40
x = sqrt(36)

xbar.minus.40 = c()

for(i in 1:mc) {
  sample = rnorm(n, mu, x)
  
  xbar.minus.40[i] = mean(sample - 40)
}

hist(xbar.minus.40, col = "blue", breaks = 50)
```

**i:**
Yes, it does look normal.
**ii.**
```{r}
mean(xbar.minus.40)
var(xbar.minus.40)
```

**Part f:**
```{r}
set.seed(1)

mc = 50000
n = 9

mu = 40
sigma = sqrt(36)

storage = c()

for(i in 1:mc) {
  sample = rnorm(n, mu, sigma)
  
  storage[i] = ((mean(sample) - mu) / (sigma/sqrt(n)))
}

hist(storage, col = "blue", breaks = 50)
```
```{r}
mean(storage)
var(storage)
```
**i.**
Yes, it is nearly $N(0,1)$.
**ii.**
Mean = -0.0001118981.
Variance = 0.9959294.

**Problem 2:**
**Part a:**
The CLT is if $x_1, ...,x_n$ are iid from a distribution with finite mean $/mu$ and finite variance $\sigma^2$, then the sampling distribution of $\bar{x}$ is approximately, for large enough n, normal.

$$\frac{\bar{x}-\mu}{\sigma/\sqrt{n}}\sim(approx)N(0,1)$$

**Part b:**
Approximately normal with sample mean $\bar{x} = \lambda$

**Part c:**
```{r}
set.seed(1)

mc = 50000

n = 40
mu = 8

storage = c()

for(i in 1:mc) {
  samp = rchisq(n, mu)
  storage[i] = mean(samp)
}

hist(storage, col = "blue", breaks = 50)
```
Yes, it looks normal.

**Part d:**
```{r}
set.seed(1)

mc = 50000

n = 4
mu = 8

storage = c()

for(i in 1:mc) {
  samp = rchisq(n, mu)
  storage[i] = mean(samp)
}

hist(storage, col = "blue", breaks = 50)
```
No because $\lambda$ can't be less than 0 due to the bounds of the chi-square distribution. So it is right skewed.

**Part e:**
Because part d has less samples so it's variance is larger.

**Part f:**
```{r}
set.seed(1)

mc = 50000

n = 40
mu = 8

storage = c()

for(i in 1:mc) {
  samp = rchisq(n, mu)
  storage[i] = mean(samp) - 8
}

hist(storage, col = "blue", breaks = 50)
```
Yes, it does look normal.

**Part g:**
```{r}
set.seed(1)

mc = 50000

n = 40
mu = 8
s = sqrt(16)

storage = c()

for(i in 1:mc) {
  samp = rchisq(n, mu)
  storage[i] = ((mean(samp) - 8)/(s/(sqrt(n))))
}

hist(storage, col = "blue", breaks = 50)

mean(storage)
var(storage)
```
The sampling distribution looks normal. The mean is -0.0006903592 and the variance is 1.008031 which are both pretty close to 0 and 1.

**Problem 3:**
**Part a:**
$s=\sqrt{\frac{\sum_{i=1}^n(x_i-\bar{x})^2}{n-1}}$

**Part b:**
```{r}
set.seed(1)

mc = 50000

n = 9
mu = 40
x = sqrt(36)

storage = c()

for(i in 1:mc) {
  samp = rnorm(n, mu, x)
  s = sqrt(sum((samp - mean(samp))^2)/(n-1))
  storage[i] = (mean(samp) - 40)/(s/sqrt(n))
}

hist(storage, col = "blue", breaks = 60)
```
Yes, it looks normal, but if we substitute in $\sigma$ for $s$, then the distribution "stops" when less than -4 and above 4. Where the t distribution "stops" at about -5 and 5.

**Part c:**
```{r}
set.seed(1)

mc = 50000

n = 20
mu = 40
x = sqrt(36)

storage = c()

for(i in 1:mc) {
  samp = rnorm(n, mu, x)
  s = sqrt(sum((samp - mean(samp))^2)/(n-1))
  storage[i] = (mean(samp) - 40)/(s/sqrt(n))
}

hist(storage, col = "blue", breaks = 50)
```

```{r}
set.seed(1)

mc = 50000

n = 30
mu = 40
x = sqrt(36)

storage = c()

for(i in 1:mc) {
  samp = rnorm(n, mu, x)
  s = sqrt(sum((samp - mean(samp))^2)/(n-1))
  storage[i] = (mean(samp) - 40)/(s/sqrt(n))
}

hist(storage, col = "blue", breaks = 50)
```
The sample standard deviation gets smaller and smaller.