---
title: "Stat 3202 Lab 8"
author: "Nathan Johnson.9254"
date: "2023-03-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Problem 1:**
**Part a:**
$(1-\alpha)\%$ conf interval for $\mu$ (when $\sigma$ is unknown) is:
$$\bar x - t_{1-\alpha/2,n-1}\frac{s}{\sqrt n} \leq \mu \leq \bar x + t_{1-\alpha/2,n-1}\frac{s}{\sqrt n}$$
**Part b:**

```{r}
set.seed(1)

mu = 10
sigma = 3
n = 5
alpha = 0.01

coverage = c()

for(i in 1:10000) {
samp = rnorm(n, mu, sigma)

xbar = mean(samp)
s = sd(samp)
t = qt(1-alpha/2, n -1)

lower = xbar - t*s/sqrt(n)
upper = xbar + t*s/sqrt(n)

coverage[i] = lower <= mu & mu <= upper
}
mean(coverage)
mean(lower)
mean(upper)
```
The coverage rate for these confidence intervals is about 99%.

**Part c:**

```{r}
set.seed(1)

mu = 10
sigma = 3
n = 50
alpha = 0.01

coverage = c()

for (i in 1:10000) {
samp = rnorm(n, mu, sigma)

output = t.test(samp, conf.level = 1-alpha)

lower = output$conf.int[1]
upper = output$conf.int[2]

coverage[i] = lower <= mu & mu <= upper
}

mean(coverage)
```
Sample size does not impact the coverage rate since the coverage rate remained 99% at n = 5 and 50.

**Part d:**
$\mu$ changed affects where the coverage is, but does not affect the distance between the lower and upper bound of the confidence interval.
Changing $\sigma$ affects how large the coverage is. A larger $\sigma$ means a larger coverage. Smaller $\sigma$ means smaller coverage.
As n gets larger, the coverage gets smaller.
As alpha gets larger, the coverage gets smaller.

**Problem 2:**
**Part a:**

```{r}
set.seed(1)
library(DescTools)

n = 40
p = 0.7

alpha = 0.05

lower = c()
upper = c()
coverage = c()

for(i in 1:10000) {
x = rbinom(1, n, p)
output = BinomCI(x, n, conf.level = 1-alpha, method = "wald")

lower[i] = output[2]
upper[i] = output[3]

coverage[i] = lower[i] <= p & p <= upper[i]
}

mean(coverage)

```
The confidence interval should have been 95%, but was 92.75% so no, coverage does not reach $(1-\alpha).$

**Part b:**

wilson: 0.9412
waldcc: 0.9597
agresti-coull: 0.9412
jeffreys: 0.9412

Waldcc is the closest. Strangely, wilson, agresti-coull, and jeffreys all came out the same.