---
title: "Stat 3202 Lab 8"
author: "Nathan Johnson.9254"
date: "2023-03-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Problem 1:**
**Part a:**
$$(\bar x - \bar y) \pm z_{1-\alpha/2} \sqrt{\frac{\sigma_x^2}{n_x}+\frac{\sigma_y^2}{n_y}}$$

**Part b:**
```{r}
set.seed(1)

mux = 10
sx = 3
nx = 5
muy = 5
sy = 5.5
ny = 8
alpha = 0.05

coverage = c()

for(i in 1:30000) {
  samplex = rnorm(nx, mux, sx)
  sampley = rnorm(ny, muy, sy)
  x = mean(samplex)
  y = mean(sampley)

  z = qnorm(1-alpha/2)

  CIlower = (x - y) - z * sqrt(sx^2/nx + sy^2/ny)
  CIUpper = (x - y) + z * sqrt(sx^2/nx + sy^2/ny)

  coverage[i] = CIlower <= mux - muy && mux - muy <= CIUpper 
}

mean(coverage)
```
The coverage rate for this confidence interval is 95% which is what we expect.

**Part c:**
**Part i.**
```{r}
set.seed(1)

mux = 10
sx = 3
nx = 5
muy = 5
sy = 3
ny = 8
alpha = 0.05

t = qt(1-alpha/2, nx+ny-2)

coverage = c()

for(i in 1:30000) {
  samplex = rnorm(nx, mux, sx)
  sampley = rnorm(ny, muy, sy)
  x = mean(samplex)
  y = mean(sampley)

  s2x = var(samplex)
  s2y = var(sampley)
  
  s2p = ((nx - 1)*s2x + (ny - 1)*s2y) / (nx + ny - 2)

  CIlower = (x - y) - t * sqrt(s2p) * sqrt(1/nx + 1/ny)
  CIUpper = (x - y) + t * sqrt(s2p) * sqrt(1/nx + 1/ny)

  coverage[i] = CIlower <= mux - muy && mux - muy <= CIUpper 
}

mean(coverage)
```
The coverage rate for this confidence interval is 95% which is what we expect.

**Part ii.**
```{r}
set.seed(1)

mux = 10
sx = 3
nx = 5
muy = 5
sy = 3
ny = 8
alpha = 0.05

coverage = c()

for(i in 1:30000) {
  samplex = rnorm(nx, mux, sx)
  sampley = rnorm(ny, muy, sy)
  
  output = t.test(samplex, sampley, conf.level = 1-alpha, var.equal = TRUE)

  CIlower = output$conf.int[1]
  CIupper = output$conf.int[2]

  coverage[i] = CIlower <= mux - muy && mux - muy <= CIUpper 
}

mean(coverage)
```
Despite copying Mr. Gan's work precisely, I can't seem to get this to run correctly. But the confidence interval should be 95%.

**Problem 2:**
**Part a:**
**Part i.**
$$v=\frac{(\frac{s_x^2}{n_x}+\frac{s_y^2}{n_y})^2}{\frac{1}{n_x-1}(\frac{s_x^2}{n_x})+\frac{1}{n_y-1}(\frac{s_y^2}{n_y})}$$
**Part ii.**
```{r}
set.seed(1)

mux = 10
sx = 3
nx = 5
muy = 5
sy = 3
ny = 8
alpha = 0.05

coverage = c()

for(i in 1:30000) {
  samplex = rnorm(nx, mean = mux, sd = sx)
  sampley = rnorm(ny, mean = muy, sd = sy)
  
  output = t.test(samplex, sampley, conf.level = 1-alpha, var.equal = FALSE)

  CIlower = output$conf.int[1]
  CIupper = output$conf.int[2]

  coverage[i] = CIlower <= (mux - muy) && (mux - muy) <= CIUpper 
}

mean(coverage)
```
Once again, it didn't seem to compute the right values, but I believe my code to be correct. The confidence interval should have a confidence of 95%.

**Problem 3:**
```{r}
set.seed(1)

mux = 10
sx = 3
nx = 5
muy = 5
sy = 3
ny = 8
alpha = 0.05

v = min(nx - 1, ny - 1)
t = qt(1-alpha/2, v)

coverage = c()

for(i in 1:30000) {
  samplex = rnorm(nx, mean = mux, sd = sx)
  sampley = rnorm(ny, mean = muy, sd = sy)
  
  x = mean(samplex)
  y = mean(sampley)
  
  CIlower = (x - y) - t * sqrt((sx^2/nx)+(sy^2/ny))
  CIupper = (x - y) + t * sqrt((sx^2/nx)+(sy^2/ny))

  coverage[i] = CIlower <= (mux - muy) && (mux - muy) <= CIUpper 
}

mean(coverage)
```
The coverage rate is 99.5% which does show that the margin of error is higher and gives a greater coverage rate than 95%.













